The Categorical Structure of Profunctor Optics

### 2017-10-30 paper.tex
Skip dual proof for prism laws

Just as lenses are constructed by two concrete maps $\fget : S \to A$ and $\fput : S \times A \to S$, prisms are constructed by two maps $\freview : A \to S$ and $\fmatching : S \to S \sqcup A$. These names are taken from the Haskell \texttt{lens} library.
\begin{align*}
\PreOptic_\sqcup(S, A) &= \int^{M \in \M} \C(S, M \sqcup A) \times \C(M \sqcup A, S) \\
&\cong \int^{M \in \M} \C(S, M \sqcup A) \times \C(M, S) \times \C(A, S) && \text{(universal property of coproduct)} \\
&\cong \C(S, S \sqcup A) \times \C(A, S) && \text{(co-Yoneda)}
\end{align*}
So, if we are given a pre-prism $\langle l, r \rangle : (S, S) \hto (A, A)$ then associated $\freview$ and $\fmatching$ morphisms are given by $\freview = r \inr$ and $\fmatching = (r\inl \sqcup \id_A)l$

The laws for prisms are the obvious duals to the lens laws. 
\begin{align*}
\fmatching \; \freview &= \inr \\
[\id_S, \freview] \fmatching &= \id_S \\
(\fmatching \sqcup \id_A) \fmatching &= \mathrm{in}_{1,3} \, \fmatching
\end{align*}

@@

\begin{proposition}
When $\C = \Set$\todo{, and probably more generally,} the third law is implied by the other two.
\end{proposition}
\begin{proof}
The key is that in $\Set$, for any map $f : X \to Y$, the set $Y$ splits into the coproduct of $\im f$ and its complement. The first law implies that $\freview$ is injective, so $S \cong C \sqcup A$ for some complement $C$. Identifying $A$ with its image in $S$, the second law implies that if $a\in A \subset S$ then $\fmatching(a) = \inr(a)$ and if $c\in C \subset S$ then $\fmatching(c) = \inl(c)$. The third law then follows easily by checking both cases.
\end{proof}

\begin{proposition}
\label{prop-OpticImpliesPrismLaws}
If $\langle l, r \rangle : S \hto A$ is a prism then the associated $\fmatching$ and $\freview$ functions satisfy the above laws.
\end{proposition}
\begin{proof}
\todo{Again it seems it is good enough if we only have $lr = t \sqcup A$}
\begin{align*}
\fmatching \; \freview 
&= (r\inl \sqcup \id_A)l r \inr \\
&= (r\inl \sqcup \id_A) \inr \\
&= \inr \\
[\id_S, \freview] \fmatching
&= [\id_S, r \inr] (r\inl \sqcup \id_A)l \\ 
&= [r\inl, r \inr] l \\ 
&= rl \\
&= \id_S
\end{align*}
\end{proof}

\subsubsection{Traversals}
\subsubsection{Setters}

>>>>

Just as lenses are constructed by two concrete maps $\fget : S \to A$ and $\fput : S \times A \to S$, prisms are constructed by two maps $\freview : A \to S$ and $\fmatching : S \to S \sqcup A$. These names are taken from the Haskell \texttt{lens} library.
\begin{align*}
\PreOptic_\sqcup(S, A) &= \int^{M \in \M} \C(S, M \sqcup A) \times \C(M \sqcup A, S) \\
&\cong \int^{M \in \M} \C(S, M \sqcup A) \times \C(M, S) \times \C(A, S) && \text{(universal property of coproduct)} \\
&\cong \C(S, S \sqcup A) \times \C(A, S) && \text{(co-Yoneda)}
\end{align*}
So, if we are given a pre-prism $\langle l, r \rangle : (S, S) \hto (A, A)$ then associated $\freview$ and $\fmatching$ morphisms are given by $\freview = r \inr$ and $\fmatching = (r\inl \sqcup A)l$

The laws for prisms are the obvious duals to the lens laws. 
\begin{align*}
\fmatching \; \freview &= \inr \\
[\id_S, \freview] \fmatching &= \id_S \\
(\fmatching \sqcup A) \fmatching &= \mathrm{in}_{1,3} \, \fmatching
\end{align*}

@@

\begin{proposition}
When $\C = \Set$\todo{, and probably more generally,} the third law is implied by the other two.
\end{proposition}
\begin{proof}
The key is that in $\Set$, for any map $f : X \to Y$, the set $Y$ is equal to the union of $\im f$ and its complement. The first law implies that $\freview$ is injective, so $S \cong C \sqcup A$ for some complement $C$. Identifying $A$ with its image in $S$, the second law implies that if $a\in A \subset S$ then $\fmatching(a) = \inr(a)$ and if $c\in C \subset S$ then $\fmatching(c) = \inl(c)$. The third law then follows easily by checking both cases.
\end{proof}

The following is then the dual of Proposition \ref{prop-OpticImpliesLensLaws}.
\begin{proposition}
\label{prop-OpticImpliesPrismLaws}
If $p : S \hto A$ is a prism then the associated $\fmatching$ and $\freview$ functions satisfy the three prism laws. \qed
\end{proposition}

\subsubsection{Traversals}
\todo{Do it}

\subsubsection{Setters}
